---
title: "Lab 1: Graphing MLS Salaries"
author: "Surabhi Damle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
knitr::opts_knit$set(
  root.dir = "."
  )
```

### 1. Load libraries and data.

```{r initialize}
# Load files, environment variables, libraries, etc. here


#install.packages("janitor")
#install.packages("Hmisc")
#install.packages("ggthemes")
#install.packages("DT")


library(tidyverse)
library(janitor)


# Source: https://s3.amazonaws.com/mlspa/2023-Salary-Report-as-of-April-30.pdf 
# Useful for liberating data from PDFs: https://tabula.technology/ 
# Context: https://mlsplayers.org/resources/salary-guide
mls_salaries <- read_csv("tabula-2023-Salary-Report-as-of-April-30.csv", 
    col_types = cols(`Base Salary` = col_number(), 
        `Calculated Guaranteed Compensation` = col_number()))


```

### 2. Doh. Looks like we have some column names that need cleaning.

```{r}
mls_salaries<- janitor::clean_names(mls_salaries)
```

### 3. Rename any of those columns that you can make shorter without making them less clear.

```{r}
names(mls_salaries) <- c("team", "last_name", "first_name", "position", "base_salary", "guaranteed_comp")
```

### 4. We always want to start by taking our data out to coffee and just getting to know it.

```{r}
mls_salaries %>% Hmisc::describe()
```

### 5. Let's arrange these by base salary.

```{r}
mls_salaries %>%
  arrange(base_salary) %>%
  DT::datatable() %>%
  DT::formatCurrency(c('base_salary', 'guaranteed_comp'),
                 digits = 0)
```

### 6. That's a lot to look at. Let's look just at players on the Charlotte's team.

```{r}
mls_salaries %>%
  filter(team == "Charlotte FC") %>%
  arrange(base_salary)
```

### 7. It looks like the players' base salaries and their guarantee compensation is often the same. I wonder what percentage of each player's guaranteed compensation is his base salary?

```{r}
mls_salaries %>%
  filter(team == "Charlotte FC") %>%
  mutate(other_compensation = guaranteed_comp - base_salary,
         pct_base = base_salary / guaranteed_comp) %>%
  arrange(desc(pct_base))
```

### 8. Looks like there's some difference there. I wonder what is "normal."

```{r}
mls_salaries %>%
  filter(team == "Charlotte FC") %>%
  mutate(other_compensation = guaranteed_comp - base_salary,
         pct_base = base_salary / guaranteed_comp) %>%
  summarise(mean(other_compensation),
            median(other_compensation))
```

### 9. I wonder how base salary differs by position across all teams.

```{r}
by_pos<- mls_salaries %>%
 group_by(position) %>%
  summarise(max_base = max(base_salary),
            median_base = median(base_salary),
            min_base = min(base_salary)
  )

by_pos
            
```

### 10. Let's see how that looks in a graph.

```{r}
date <- "April 30, 2023"

ggplot(data = by_pos,
       mapping = aes(x = fct_reorder(position,
                         median_base), 
                     y = median_base)) +
  geom_col() +
  scale_x_discrete(labels=c("GK" = "Goalkeepers", 
                            "D" = "Defenders",
                            "M" = "Midfielders",
                            "F" = "Forwards")) +
  # show y axis in dollar format
  scale_y_continuous(labels=scales::dollar_format()) +
  # use minimal theme
  ggthemes::theme_fivethirtyeight() +
  # customize labels
  labs(
    x = "Player Positions",
    y = "Median Base Salary",
    title = "Median base salary for MLS players",
    subtitle = paste0("Data as of ", date, .sep = " "),
    caption = "Source: mlsplayers.org/resources/salary-guide"
  )
```

### 11. Let's save that as a file we can use elsewhere.

```{r}
my_chart <- ggplot(data = by_pos,
       mapping = aes(x = fct_reorder(position,
                         median_base), 
                     y = median_base)) +
  geom_col() +
  scale_x_discrete(labels=c("GK" = "Goalkeepers", 
                            "D" = "Defenders",
                            "M" = "Midfielders",
                            "F" = "Forwards")) +
  # show y axis in dollar format
  scale_y_continuous(labels=scales::dollar_format()) +
  # use minimal theme
  ggthemes::theme_fivethirtyeight() +
  # customize labels
  labs(
    x = "Player Positions",
    y = "Avg. Base Salary",
    title = "Median base salary for MLS players",
    subtitle = paste0("Data as of", date, .sep = " "),
    caption = "Source: mlsplayers.org/resources/salary-guide"
  )               

ggsave(file="salaries_by_pos.jpg",
       device = "jpg",
       plot = my_chart,
       path = NULL
       )
```


```{r}
by_team<-mls_salaries %>%
 group_by(team) %>%
  summarise(max_base = max(base_salary),
            average_base = mean(base_salary),
            min_base = min(base_salary),
            median_base=
median(base_salary)
  )

by_team%>%
  DT::datatable() %>%
  DT::formatCurrency(c('max_base', 'average_base','min_base', 'median_base'),
                 digits = 0)
```
